---
title: "Merged Metrics"
author: "Sarah Speroff"
date: "3/1/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# Experimental data details

Colony:	Either A (for aposymbiotc) or S (for symbiotic) and in ascending order. Counts the number of colonies of each phenotype used														
Fragment:	The # fragment that was excised from a given colony														
BTag:	Identification code that differentiates every coral polyp from the rest, letter represents the color of the tag														
Phenotype:	either aposymbiotic or symbiotic polyp as determined from the colony of origin														
Treatment:	Treatment codes differentiate the unique combination of stressors that make up that experimental condition: Plastic or Noplastic, Light or Dark, Fed or Starved														
Temperature:	Either ambient (19 degrees C) or elevated (30 degrees C)														
Tank:	Tank location on X-Rack (row/column) or wetlab tray														
Tile:	Unique acrylic disk that held 5 polpys per tank (all same phenotype)														
DaysAlive:	The total number of days the fragment survived during the 22 day experiment														
Censored:	Y/N: Did we lose that polyp to a reason outside of mortality? (ie. did it get lost, or crushed during removal for respiration?)														
Mass.day0: Polpy dry(ish) mass in Grams (G) on the day that the polp was fragmented from the colony (post-frag)														
Mass.day8:	Polpy dry(ish) mass in Grams on day 8														
Mass.day15:	Polpy dry(ish) mass in Grams on day 15														
Mass.day22:	Polpy dry(ish) mass in Grams on day 22														
FvFm.day1:	Photosynthetic efficiency of coral symbiots measured using PAM on day of the experiment after ~30 minutes dark acclimated														
FvFm.day8:	" " day 8														
FvFm.day15:	" " day 15														
FvFm.day22:	" " day 22														
CoralR.day0:	Red color channel intensity corrected to background red standard using photographs as a non-destructive proxy for chlorophyll-a density. Collected on the day the polyp was fragmented from the colony (post-frag)														
CoralR.day1:	" " on the first day of experimental treatments (pre-treat)														
CoralR.day22:	" " on the last day of experimental treatments (post-treat)														
Bacteria:	Bacterial challenge treatment: Either lysed E.coli cells (EC) or saltwater controls (SW)														
MR.day8:	Metabolic rate (oxygen consumption) in umols O2 per minute per gram of polyp mass on day 8. Values are corrected to polyp mass and respirometer chamber volume	MR.day15:	MR day 15														
MR.day22:	MR day 22														
```

###########################################################################################################################
#
#                                              Data Prep, and transformations
#
############################################################################################################################

```{r}
library(dplyr)
library(tidyr)
```

```{r}
# set working directory
setwd("/Users/speroffs/Desktop/A.Poculata-Multiple-stressors-thesis-analysis-")
```
```{r}
# Load dataset with merged metrics 
data <- read.csv("All.data.merged.csv", header=TRUE)
```
```{r}
# Change data types 
data$Colony <- as.factor(data$Colony)
data$BTag <- as.factor(data$BTag)
data$Phenotype <- as.factor(data$Phenotype)
data$Treatment <- as.factor(data$Treatment)
data$Temperature <- as.factor(data$Temperature)
data$Tank <- as.factor(data$Tank)
data$Bacteria <- as.factor(data$Bacteria)
data$DaysAlive <- as.numeric(data$DaysAlive)
```

# Let's transform our dataset from it's wide format to a long format
```{r}
# Polpy ID data
meta.data <- data %>%
  select(Colony:Censored, Bacteria)
```
```{r}
# Mass data
mass.long.data <- data %>%
  select(BTag, MassG.day0:MassG.day22) %>%
  rename(c(Day0 = MassG.day0, Day8 = MassG.day8, Day15 = MassG.day15, Day22 = MassG.day22)) %>%
  pivot_longer(
    cols = c("Day0", "Day8", "Day15", "Day22"),
    names_to = "Day",
    values_to = "Mass.G") 
```
```{r}
# Photosynthetic Efficiency data
fvfm.long.data <- data %>%
  select(BTag, FvFm.day1:FvFm.day22) %>%
  rename(c(Day1 = FvFm.day1, Day8 = FvFm.day8, Day15 = FvFm.day15, Day22 = FvFm.day22)) %>%
  pivot_longer(
    cols = c("Day1", "Day8", "Day15", "Day22"),
    names_to = "Day",
    values_to = "Fvfm") 
```
```{r}
# Color data
rgb.long.data <- data %>%
  select(BTag, CoralR.day0:CoralR.day22) %>%
  rename(c(Day0 = CoralR.day0, Day1 = CoralR.day1, Day22 = CoralR.day22)) %>%
  pivot_longer(
    cols = c("Day0", "Day1", "Day22"),
    names_to = "Day",
    values_to = "Rcolor") 
```
```{r}
# Respiration data
mr.long.data <- data %>%
  select(BTag, MR.day8:MR.day22) %>%
  rename(c(Day8 = MR.day8, Day15 = MR.day15, Day22 = MR.day22)) %>%
  pivot_longer(
    cols = c("Day8", "Day15", "Day22"),
    names_to = "Day",
    values_to = "MR") 
```

# Now that we have the long format for each of our variables, let's transform our Metabolic rate and Mass data to before we join them all back into one dataframe
```{r}
# We know that from our previous work with the respiration data that a log-transformation is needed, but let's take a look to make sure
hist(mr.long.data$MR)

mr.long.data$MR <-abs(mr.long.data$MR) # we'll start by taking the absolute value of our MR data to attribute a higher number to a faster rate
mr.long.data$logMR <- log(mr.long.data$MR) # and now we can log-transform that value

hist(mr.long.data$logMR) # a beautiful histogram! This is the data we want to bring back into our larger dataset
```
```{r}
# Now let's look at our Mass data 
hist(mass.long.data$Mass.G) # not the ideal skew

# for the sake of continuity, we will convert the mass into milligrams like we did in our growth analysis
mass.long.data$mass.mg <- mass.long.data$Mass.G*1000
hist(mass.long.data$mass.mg) # does not change the skew, so let's log transform like we did last time

mass.long.data$logmass.mg <- log(mass.long.data$mass.mg)
hist(mass.long.data$logmass.mg) # much better, we'll work with this data
```

# Now, lets bring all of those long formatted datasets back together with our log-transformed mass and metabolic rate data
```{r}
all.data <- 
  merge(
    merge(
      merge(
        merge(
          mass.long.data,
        fvfm.long.data, all = TRUE),
      rgb.long.data, all = TRUE),
    mr.long.data, all = TRUE),
  meta.data,all = TRUE)

```
```{r}
# Save this long dataframe as a csv.file so we don't have to do this every time!
path <- "/Users/speroffs/Desktop/A.Poculata-Multiple-stressors-thesis-analysis-"
write.csv(all.data, file.path(path, "All.data.merged.long.format.for.R.csv"))
```

###########################################################################################################################
#
#                                                  Start here!
#
############################################################################################################################

```{r}
library(ggplot2)
```

```{r}
# Let's load in that merged dataframe!
merge.data <- read.csv("All.data.merged.long.format.for.R.csv", header=TRUE)
```

```{r}
ggplot(merge.data,
       aes(x = DaysAlive,
           y = logmass.mg,
           color = Treatment,
           group = Treatment)) +
         geom_point() +
  geom_line() +
  theme_classic()
   
```
```{r}
# Let's write a function to bin our survival data
If an individual (BTag) in merge.data has a DaysAlive value < 8, put it in a new column (survbin) and call it 8
if an individual in merge data has a days alive value between 8 and 15, put it in survbin and call it 15
if an individual in merge data has a days alive value > 15, put it in surv bin and call it 22
```
```{r}
# set up cut-off values 
breaks <- c(0,8,15,22)
# specify interval/bin labels
tags <- c("0","1-7","8-14", "15-22")

test <- merge.data %>% 
  select(BTag, Treatment, DaysAlive) 
survgroup <- as_tibble(test) %>% 
  mutate(survbin = case_when(
    DaysAlive <= 8 ~ tags[1],
    DaysAlive > 8 & DaysAlive <= 15 ~ tags[2],
    DaysAlive > 15 & DaysAlive <= 22 ~ tags[3],
    ))

survgroup$survbin <- factor(survgroup$survbin,
                       levels = tags,
                       ordered = FALSE)
summary(survgroup$survbin)

duplicated(survgroup)
survgroup = survgroup %>% 
  unique()
survgroup
```
```{r}
# plot the number of individuals in 
ggplot(survgroup, mapping = aes(x=survbin, color = Treatment, fill = Treatment)) +
  geom_bar() +
  facet_wrap(vars(Treatment)) +
  theme_classic()
## SHOOT, they are located in the bins in which they died. I want them to be in the bins in which they lived
```

```{r}
# HOW can i bin the data so that I can plot mr/mass/growth/etc and SURVIVAL over time and still retain the data points at their respective days?
test2 <- merge.data %>% 
  select(BTag, Treatment, DaysAlive, Day) 
survgroup2 <- as_tibble(test2) %>% 
  mutate(survbin = case_when(
    Day = "Day0"~ tags[1],
    DaysAlive < 8 ~ tags[2],
    DaysAlive >= 8 & DaysAlive <= 14 ~ tags[3],
    DaysAlive >= 15 & DaysAlive <= 22 ~ tags[4]
    ))
```










