---
title: "Survival.RMD"
author: "Sarah Speroff"
date: "11/18/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

knitr::opts_chunk$set(
  fig.path = "images/")

knitr::opts_chunk$set(dev = "png",
                      dpi = 300)
```

######################################################################################################
#
#                                               Data prep
#
######################################################################################################

```{r library, message = FALSE}
library(survival)
library(ggplot2)
library(dplyr)
library(ggfortify)
library(survminer)
library(RColorBrewer)
library(ggthemes)
```

```{r wd}
setwd("/Users/speroffs/Desktop/A.Poculata-Multiple-stressors-thesis-analysis-/Survival")
```
```{r load data}
# Load data for GRAPHICAL PARAMETERS 
graph.dat <- read.csv("Full_ daily mortality tracker - Censored.CoxPH.w.controls.csv", 
                     header = TRUE) 

# Load data for MODEL CREATION (1 death added to each treatment to prevent infinite coeff)
surv.dat <- read.csv("2Full_ daily mortality tracker - Censored.CoxPH.w.controls.csv", 
                     header = TRUE) 
```

```{r relevel factors}
# relevel to compare all to control 
surv.dat$Treatment <- factor(surv.dat$Treatment, levels = c("amb.ctrl", "PDS", "PLS", "NDS", "NLS", "PDF", "PLF", "NDF", "NLF"))

graph.dat$Treatment <- factor(graph.dat$Treatment, levels = c("amb.ctrl", "PDS", "PLS", "NDS", "NLS", "PDF", "PLF", "NDF", "NLF"))
```

```{r as factor}
# as factor for df 1
surv.dat$Plastic <- as.factor(surv.dat$Plastic)
surv.dat$Light <- as.factor(surv.dat$Light)
surv.dat$Food <- as.factor(surv.dat$Food)
surv.dat$Treatment <- as.factor(surv.dat$Treatment)
surv.dat$Sym.State <- as.factor(surv.dat$Sym.State)

# for df 2
graph.dat$Plastic <- as.factor(graph.dat$Plastic)
graph.dat$Light <- as.factor(graph.dat$Light)
graph.dat$Food <- as.factor(graph.dat$Food)
graph.dat$Treatment <- as.factor(graph.dat$Treatment)
graph.dat$Sym.State <- as.factor(graph.dat$Sym.State)
```

######################################################################################################
#
#          Kaplan-Meier Survival Function for all experimental treatments and control polyps
#
######################################################################################################

```{r treat.mod}
# Create a survival model to assess mortality over time by treatment with modeling data
treat.mod <- coxph(Surv(Time, Death) ~Treatment, data=surv.dat)

# Summarize the model 
summary(treat.mod)

## All experimental treatments experienced significanlty higher rates of mortality compared to ambient conditions (reference group) # to change reference: re-level treatment order 
```

```{r treat.mod LR test}
# Perform Log-rank test on our treat.mod
treat_diff <- survdiff(Surv(Time, Death) ~ Treatment, data=surv.dat)
treat_diff
```

```{r KM.survival.plot.all.treatments}
# Write the same model using the graphing data, and plot that 
treat <- survfit(Surv(Time, Death) ~Treatment, data=graph.dat)

# Plot kaplan meier survival function with ggplot
ggsurvplot(treat, # survival function
                conf.int = FALSE, # remove confidence intervals
                pval = FALSE, # do not include p value on plot
                surv.scale = "default", # or %
                legend.labs = c("Control", "PDS", "PLS", "NDS", "NLS", "PDF", "PLF", "NDF", "NLF"), 
                legend = c(0.1, 0.4), font.legend = 14, legend.title ="", # legend parameters 
                xlab = "Day", xlim = c(0, 23), break.x.by = 2, font.x = 14, # x axis parameters
                ylim = c(0.65,1), font.y= 14, # y axis parameters
                font.tickslab = 14,
                censor.size = 6,
                ggtheme = theme_classic2(),
                title = "A",
                palette = c("#131513","#F53100", "#FF8B47","#C39809","#F6CB3C",
                            "#0B8454", "#7BB12F","#3F88C5","#A33461"))

```

##############################################################################################################
#
#                  KM Survival Functions and Cox Proportional Hazards for individual stressors
#
##############################################################################################################

# Isolate experimental treatments in a new data frame to look at the risk of plastic, food, and light stress on mortality at elevated temperature 

```{r subset experimental data}
# Subset data for statistical analysis 
# Can use the "graph.dat" dataframe because we are not comparing to ambient control polyps (which did not exprience mortality)
exp.data <- filter(graph.dat, Treatment != "amb.ctrl")
summary(exp.data) # 40 indiv per treat, 160 per stress
```

###########
# PLASTIC #   
###########

## Start with looking at the effect of plastic including ASW conditions
```{r plastic all data}
# CPH model to assess the mortality risk associated with exposure to plastics 
plastic.mod <- coxph(Surv(Time, Death) ~ Plastic, data=surv.dat)
summary(plastic.mod)
```
```{r plastic all data LR}
# Perform Log-rank test on our plastic.mod
plastic_diff <- survdiff(Surv(Time, Death) ~ Plastic, data=surv.dat)
plastic_diff
```
```{r KM.plastic.all.data}
# Write the same model as a survial function
plastic <- survfit(Surv(Time, Death) ~ Plastic, data = surv.dat)

# Plot
ggsurvplot(plastic, 
          conf.int = FALSE, pval = FALSE, surv.scale = "default", # or %
          legend.labs = c("No plastic", "Plastic"), 
          legend.title = "", legend = c(0.2, 0.2), font.legend = 20,
          xlab = "", xlim = c(0, 23),break.x.by = 2, font.x = 14,
          ylab = "", ylim = c(0.7,1),font.y= 14,
          font.tickslab = 14,
          censor.size = 6,
          ggtheme = theme_classic2(),
          linetype = c(1,1),
          palette = c("black","forestgreen")) 

```

## Now subset to analyze within experimental conditions
```{r plasticCPH.experimental}
# CPH model
plastic.mod1 <- coxph(Surv(Time, Death) ~ Plastic, data=exp.data)
summary(plastic.mod1)
```
```{r EXP.KM.PLASTIC}
# KM function for plastic model
plastic1 <- survfit(Surv(Time, Death) ~ Plastic, data = exp.data)

# plot
ggsurvplot(plastic1, 
         conf.int = FALSE, # no confidence intervals 
          # pval = TRUE,  # show p value, default as numeric
          pval = "LRT: p = 0.2", # custom p value string, from likelihood ratio test
          pval.coord = c(15,1), 
          pval.size = 6, # default size is 5
          surv.scale = "default", # or %
          legend.labs = c("No Plastic", "Plastic"), 
          legend.title = "", legend = c(0.2, 0.2), font.legend = 20,
          xlab = "", xlim = c(0, 23),break.x.by = 2, font.x = 14,
          ylab = "", ylim = c(0.7,1),font.y= 14,
          font.tickslab = 14,
          censor.size = 6,
          ggtheme = theme_classic2(),
          title = "B. Plastic", 
          font.title = c(20, "bold"),
          linetype = c(1,1),
          palette = c("black","forestgreen")) 
```

########
# FOOD #   
########

## Start with looking at the effect of feeding including ASW conditions
```{r}
# CPH model
food.mod <- coxph(Surv(Time, Death) ~ Food, data=surv.dat)
summary(food.mod)
```
```{r}
# Perform Log-rank test on our food.mod
food_diff <- survdiff(Surv(Time, Death) ~ Food, data=surv.dat)
food_diff
```
```{r}
# Kaplan meier curve for food mod
food <- survfit(Surv(Time, Death) ~ Food, data = surv.dat)
```
```{r}
# plot 
ggsurvplot(food, 
          conf.int = FALSE, pval = FALSE, surv.scale = "default", # or %
          legend.labs = c("Fed", "Starved"), 
          legend.title = "", legend = c(0.2, 0.2), font.legend = 20,
          xlab = "", xlim = c(0, 23),break.x.by = 2, font.x = 14,
          ylab = "", ylim = c(0.75,1),font.y= 14,
          font.tickslab = 14,
          censor.size = 6,
          ggtheme = theme_classic2(),
          linetype = c(1,1),
          palette = c("orangered4","black")) 
```

## Now subset to analyze within experimental conditions
```{r}
# CPH model
food.mod1 <- coxph(Surv(Time, Death) ~ Food, data=exp.data)
summary(food.mod1)
```
```{r EXP.KM.FOOD}
# KM for food model 1
food1 <- survfit(Surv(Time, Death) ~ Food, data = exp.data)

# plot
ggsurvplot(food1, 
          conf.int = FALSE, # no confidence intervals 
          # pval = TRUE,  # show p value, default as numeric
          pval = "LRT: p < 0.001***", # custom p value string, from likelihood ratio test
          pval.coord = c(15,1), 
          pval.size = 6, # default size is 5
          surv.scale = "default", # or %
          legend.labs = c("Fed", "Starved"), 
          legend.title = "", legend = c(0.2, 0.2), font.legend = 20,
          xlab = "", xlim = c(0, 23),break.x.by = 2, font.x = 14,
          ylab = "", ylim = c(0.7,1),font.y= 14,
          font.tickslab = 14,
          censor.size = 6,
          ggtheme = theme_classic2(),
          title = "D. Food", 
          font.title = c(20, "bold"),
          linetype = c(1,1),
          palette = c("orangered4","black")) 
```

########
# FOOD: Fed vs Starved in No Plastic / Light treatment   
########

```{r}
# Filter out observations: with just Plastic/Dark/Fed variables 
NL.dat <- filter(surv.dat, Treatment == "NLF" | Treatment == "NLS")
```
```{r}
# CPH model
nl.cox <- coxph(Surv(Time, Death) ~ Food, data=NL.dat)
summary(nl.cox)

## Loglikelihoof converged: infinite coefficient!
```
```{r}
# Perform Log-rank test 
nl_diff <- survdiff(Surv(Time, Death) ~ Food, data=NL.dat)
nl_diff
```
```{r}
# KM function to plot
nl <- survfit(Surv(Time, Death) ~ Food, data = NL.dat)

# plot 
ggsurvplot(nl, conf.int = FALSE, 
           pval = FALSE, 
           surv.scale = "default", # or %
           legend.labs = c("Fed", "Starved"),
           legend.title = "",
           legend = c(0.15, 0.2),
           font.legend = 10,
           xlab = "Days",
           xlim = c(0, 23),
           break.x.by = 2,
           font.x = 12,
           ylim = c(0.6,1),
           font.y= 12, 
           censor.size = 5,
           ggtheme = theme_classic2(),
           palette = c("orangered4", "black"),
           title = "No Plastic / Light")
```

########
# FOOD: Fed vs Starved in Plastic / Dark treatment   
########

```{r}
# filter out observations: with just Plastic/Dark variables 
PD.dat <- filter(surv.dat, Treatment == "PDF" | Treatment == "PDS")
```
```{r}
# CPH model
PD.cox <- coxph(Surv(Time, Death) ~ Food, data=PD.dat)
summary(PD.cox)
```
```{r}
# Perform Log-rank test 
PD_diff <- survdiff(Surv(Time, Death) ~ Food, data=PD.dat)
PD_diff
```
```{r}
# KM function 
PD <- survfit(Surv(Time, Death) ~ Food, data = PD.dat)

# plot 
ggsurvplot(PD, conf.int = FALSE, 
           pval = FALSE, 
           surv.scale = "default", # or %
           legend.labs = c("Fed", "Starved"),
           legend.title = "",
           legend = c(0.15, 0.2),
           font.legend = 10,
           xlab = "Days",
           xlim = c(0, 23),
           break.x.by = 2,
           font.x = 12,
           ylim = c(0.6,1),
           font.y= 12, 
           censor.size = 5,
           ggtheme = theme_classic2(),
           palette = c("orangered4", "black"),
           title = "Plastic / Dark")
```

########
# FOOD: Fed vs Starved in No Plastic / Dark treatment   
########

```{r}
# filter out observations with just No Plastic/Dark variables 
ND.dat <- filter(surv.dat, Treatment == "NDF" | Treatment == "NDS")
```
```{r}
# CPH model
ND.cox <- coxph(Surv(Time, Death) ~ Food, data=ND.dat)
summary(ND.cox)
```
```{r}
# Perform Log-rank test 
ND_diff <- survdiff(Surv(Time, Death) ~ Food, data=ND.dat)
ND_diff
```
```{r}
# KM function
ND <- survfit(Surv(Time, Death) ~ Food, data = ND.dat)

# plot 
ggsurvplot(ND, conf.int = FALSE, 
           pval = FALSE, 
           surv.scale = "default", # or %
           legend.labs = c("Fed", "Starved"),
           legend.title = "",
           legend = c(0.15, 0.2),
           font.legend = 10,
           xlab = "Days",
           xlim = c(0, 23),
           break.x.by = 2,
           font.x = 12,
           ylim = c(0.6,1),
           font.y= 12, 
           censor.size = 5,
           ggtheme = theme_classic2(),
           palette = c("orangered4", "black"),
           title = "No Plastic / Dark")
```

########
# FOOD: Fed vs Starved in Plastic / Light treatment   
########

```{r}
# filter out observations with just Plastic/Light variables 
PL.dat <- filter(surv.dat, Treatment == "PLF" | Treatment == "PLS")
```
```{r}
# CPH model
PL.cox <- coxph(Surv(Time, Death) ~ Food, data=PL.dat)
summary(PL.cox)
```
```{r}
# Perform Log-rank test 
PL_diff <- survdiff(Surv(Time, Death) ~ Food, data=PL.dat)
PL_diff
```
```{r}
# KM function
PL <- survfit(Surv(Time, Death) ~ Food, data = PL.dat)

# plot 
ggsurvplot(PL, conf.int = FALSE, 
           pval = FALSE, 
           surv.scale = "default", # or %
           legend.labs = c("Fed", "Starved"),
           legend.title = "",
           legend = c(0.15, 0.2),
           font.legend = 10,
           xlab = "Days",
           xlim = c(0, 23),
           break.x.by = 2,
           font.x = 12,
           ylim = c(0.6,1),
           font.y= 12, 
           censor.size = 5,
           ggtheme = theme_classic2(),
           palette = c("orangered4", "black"),
           title = "Plastic / Light")
```


#### New code (Speroff, 11/16/2022): The effects of plastic under starvation:
```{r}
# Subset data to isolate all starved polyps
st.dat <- filter(surv.dat, Treatment == "PDS" | Treatment == "PLS" | 
                  Treatment == "NDS"| Treatment == "NLS")
```
```{r}
# CPH model
pl_st.cox <- coxph(Surv(Time, Death) ~ Plastic, data=st.dat)
summary(pl_st.cox)
```
```{r}
# Perform Log-rank test 
pl_st_diff <- survdiff(Surv(Time, Death) ~ Plastic, data=st.dat)
pl_st_diff
```
```{r}
# KM function
pl_st <- survfit(Surv(Time, Death) ~ Plastic, data = st.dat)

# plot 
ggsurvplot(pl_st, 
          conf.int = FALSE, # no confidence intervals 
          # pval = TRUE,  # show p value, default as numeric
          pval = "LRT: p = 0.005*", # custom p value string, from likelihood ratio test
          pval.coord = c(16,0.72), 
          pval.size = 6, # default size is 5
          surv.scale = "default", # or %
          legend.labs = c("No Plastic", "Plastic"), 
          legend.title = "", legend = c(0.2, 0.2), font.legend = 20,
          xlab = "Survival Probability", xlim = c(0, 23),break.x.by = 2, font.x = 14,
          ylab = "Days", ylim = c(0.7,1),font.y= 14,
          font.tickslab = 14,
          censor.size = 6,
          ggtheme = theme_classic2(),
          title = "Effect of plastic exposure during starvation", 
          font.title = c(20, "bold"),
          linetype = c(1,1),
          palette = c("black","darkgreen")) 
```

#########
# Light #   
#########

## Start with looking at the effect of feeding including ASW conditions
```{r}
# Light
light.mod <- coxph(Surv(Time, Death) ~ Light, data=surv.dat)
summary(light.mod)
```
```{r}
# Perform Log-rank test on our light.mod
light_diff <- survdiff(Surv(Time, Death) ~ Light, data=surv.dat)
light_diff
```
```{r}
# kaplan meier curve 
light <- survfit(Surv(Time, Death) ~ Light, data = graph.dat)

# plot 
ggsurvplot(light, 
         conf.int = FALSE, pval = FALSE, surv.scale = "default", # or %
          legend.labs = c("Dark", "Light"), 
          legend.title = "", legend = c(0.2, 0.2), font.legend = 20,
          xlab = "", xlim = c(0, 23),break.x.by = 2, font.x = 14,
          ylab = "", ylim = c(0.75,1),font.y= 14,
          font.tickslab = 14,
          censor.size = 6,
          ggtheme = theme_classic2(),
          linetype = c(1,1),
          palette = c("black","darkgoldenrod1")) 
```

## Now subset to analyze within experimental conditions
```{r}
# CPH model
light.mod1 <- coxph(Surv(Time, Death) ~ Light, data=exp.data)
summary(light.mod1)
```
```{r EXP.KM.LIGHT}
# KM function
light1 <- survfit(Surv(Time, Death) ~ Light, data = exp.data)

# plot
ggsurvplot(light1, 
         conf.int = FALSE, # no confidence intervals 
          # pval = TRUE,  # show p value, default as numeric
          pval = "LRT: p = 0.7", # custom p value string, from likelihood ratio test
          pval.coord = c(15,1), 
          pval.size = 6, # default size is 5
          surv.scale = "default", # or %
          legend.labs = c("Dark", "Light"), 
          legend.title = "", legend = c(0.2, 0.2), font.legend = 20,
          xlab = "", xlim = c(0, 23),break.x.by = 2, font.x = 14,
          ylab = "", ylim = c(0.7,1),font.y= 14,
          font.tickslab = 14,
          censor.size = 6,
          ggtheme = theme_classic2(),
          title = "C. Light", 
          font.title = c(20, "bold"),
          linetype = c(1,1),
          palette = c("black","darkgoldenrod1")) 
```

#########
# Light: Apo frags
#########

```{r}
# Filter
apo.dat <- filter(surv.dat, Sym.State == "Apo")
```

```{r}
# cox model for apo
apo.coxm <- coxph(Surv(Time, Death) ~ Light, data=apo.dat)
summary(apo.coxm)
```
```{r}
# Perform Log-rank test 
apo_diff <- survdiff(Surv(Time, Death) ~ Light, data=apo.dat)
apo_diff
```

```{r}
# surv fit 
afit <- survfit(Surv(Time, Death) ~ Light, data = apo.dat)
# plot 
a.plot <- ggsurvplot(afit, conf.int = FALSE, 
           pval = FALSE, 
           surv.scale = "default", # or %
           legend.labs = c("Dark", "Light"),
           legend.title = "",
           legend = c(0.15, 0.2),
           font.legend = 10,
           xlab = "Days",
           xlim = c(0, 23),
           break.x.by = 2,
           font.x = 12,
           ylim = c(0.6,1),
           font.y= 12, 
           censor.size = 5,
           ggtheme = theme_classic2(),
           palette = c("black", "goldenrod3"),
           title = "Aposymbiotic")

a.plot
```

#########
# Light: Sym frags
#########

```{r}
# Filter
sym.dat <- filter(surv.dat, Sym.State == "Sym")
```
```{r}
# cox model for sym
sym.coxm <- coxph(Surv(Time, Death) ~ Light, data=sym.dat)
summary(sym.coxm)
```
```{r}
# Perform Log-rank test 
sym_diff <- survdiff(Surv(Time, Death) ~ Light, data=sym.dat)
sym_diff
```
```{r}
# surv fit 
sfit <- survfit(Surv(Time, Death) ~ Light, data = sym.dat)
# plot 
ggsurvplot(sfit, conf.int = FALSE, 
           pval = FALSE, 
           surv.scale = "default", # or %
           legend.labs = c("Dark", "Light"),
           legend.title = "",
           legend = c(0.15, 0.2),
           font.legend = 10,
           xlab = "Days",
           xlim = c(0, 23),
           break.x.by = 2,
           font.x = 12,
           ylim = c(0.6,1),
           font.y= 12, 
           censor.size = 5,
           ggtheme = theme_classic2(),
           palette = c("black", "goldenrod3"),
           title = "Symbiotic")
```

###############
# Temperature    
###############

```{r}
# Temperature
temperature.mod <- coxph(Surv(Time, Death) ~ Temperature, data=surv.dat)
summary(temperature.mod)
```
```{r}
# Perform Log-rank test on our temperature.mod
temperature_diff <- survdiff(Surv(Time, Death) ~ Temperature, data=surv.dat)
temperature_diff
```
```{r KM TEMP}
# kaplan meier curve 
temperature <- survfit(Surv(Time, Death) ~ Temperature, data = graph.dat)
# plot 
ggsurvplot(temperature, 
         conf.int = FALSE, # no confidence intervals 
          # pval = TRUE,  # show p value, default as numeric
          pval = "LRT: p < 0.001***", # custom p value string, from likelihood ratio test
          pval.coord = c(15,0.95), 
          pval.size = 6, # default size is 5
          surv.scale = "default", # or %
          legend.labs = c("Ambient (19 C)", "Elevated (30 C)"), 
          legend.title = "", legend = c(0.3, 0.2), font.legend=20,
          xlab = "", xlim = c(0, 23),break.x.by = 2, font.x = 14,
          ylab = "", ylim = c(0.7,1),font.y= 14,
          font.tickslab = 14,
          censor.size = 6,
          ggtheme = theme_classic2(),
          title = "A. Temperature", 
          font.title = c(20, "bold"),
          linetype = c(1,1),
          palette = c("royalblue4","firebrick2")) 
```

###############
# Temperature: ambient vs elevated controls 
###############

```{r}
# Ambient vs elevated controls
# filter out observations with just amb
tmp.dat <- filter(surv.dat, Treatment == "NLF" | Treatment == "amb.ctrl")
```

```{r}
# cox model for temperature controls
temperature.mod1 <- coxph(Surv(Time, Death) ~ Temperature, data=tmp.dat)
summary(temperature.mod1)
```
```{r}
# Perform Log-rank test 
tmp_diff <- survdiff(Surv(Time, Death) ~ Temperature, data=tmp.dat)
tmp_diff
```

```{r}
# surv fit 
temperature1 <- survfit(Surv(Time, Death) ~ Temperature, data = tmp.dat)
# plot 
ggsurvplot(temperature1, 
           conf.int = FALSE, 
           pval = FALSE, 
           surv.scale = "default", # or %
           legend.labs = c("Ambient (19 C)", "Elevated (30 C)"),
           legend.title = "",
           legend = c(0.15, 0.2),
           font.legend = 10,
           xlab = "Days",
           xlim = c(0, 23),
           break.x.by = 2,
           font.x = 12,
           ylim = c(0.6,1),
           font.y= 12, 
           censor.size = 5,
           ggtheme = theme_classic2(),
           palette = c("darkblue", "red"),
           title = "Temperature Controls")

```

#############
# Phenotype   
#############

```{r}
# Sym.State
phen.mod <- coxph(Surv(Time, Death) ~ Sym.State, data=surv.dat)
summary(phen.mod)
```
```{r}
# Perform Log-rank test on our temperature.mod
phen_diff <- survdiff(Surv(Time, Death) ~ Sym.State, data=surv.dat)
sym_diff
```
```{r}
# surv fit 
phen <- survfit(Surv(Time, Death) ~ Sym.State, data = graph.dat)
# plot 
ggsurvplot(phen, 
          conf.int = FALSE, pval = FALSE, surv.scale = "default", # or %
          legend.labs = c("Aposymbiotic", "Symbiotic"), 
          legend.title = "", legend = c(0.2, 0.2), font.legend = 16,
          xlab = "Experimental Day", xlim = c(0, 23),break.x.by = 2, font.x = 14,
          ylab = "Survival Probability", ylim = c(0.75,1),font.y= 14,
          title = "Symbiotic Phenotype", font.title = c(16,"bold"),
          font.tickslab = 14,
          censor.size = 6,
          ggtheme = theme_classic2(),
          linetype = c(1,1),
          palette = c("tan","coral4"))


```

## Phenotype in experimental conditions
```{r}
# CPH model
phen.mod1<- coxph(Surv(Time, Death) ~ Sym.State, data=exp.data)
summary(phen.mod1)
```
```{r EXP PHENOTYPE}
# plot graph data (also subsetted for elevated temp)
phen1 <- survfit(Surv(Time, Death) ~ Sym.State, data = exp.data)

# plot
ggsurvplot(phen1, 
          conf.int = FALSE, # no confidence intervals 
          # pval = TRUE,  # show p value, default as numeric
          pval = "Schoenfeld's Residuals: p < 0.01**", # custom p value string, from likelihood ratio test
          pval.coord = c(13.5,0.98), 
          pval.size = 4, # default size is 5
          surv.scale = "default", # or %
          legend.labs = c("Aposymbiotic", "Symbiotic"), 
          legend.title = "", legend = c(0.2, 0.2), font.legend=14,
          xlab = "Day", xlim = c(0, 23),break.x.by = 2, font.x = 14,
          ylab = "Survival Probability", ylim = c(0.7,1),font.y= 14,
          font.tickslab = 14,
          censor.size = 6,
          ggtheme = theme_classic2(),
          title = "Symbiotic Phenotype", 
          font.title = c(20, "bold"),
          linetype = c(1,1),
          palette = c("tan","coral4"))
```

# Run single variable CPH models simultaneously, for plastic, food, light, and phenotype within the elevated temperature experimental treatment

```{r UNIVARIATE}
# String of factors 
covariates <- c("Plastic", "Food",  "Light", "Sym.State")

# Formula for the CPH model
exp_formulas <- sapply(covariates,
                        function(x) as.formula(paste('Surv(Time, Death)~', x)))
# # Run the models                        
exp_models <- lapply(exp_formulas, function(x){coxph(x, data = exp.data)})

# Extract model output 
exp_results <- lapply(exp_models,
                       function(x){ 
                          x <- summary(x)
                          p.value<-signif(x$wald["pvalue"], digits=2)
                          wald.test<-signif(x$wald["test"], digits=2)
                          beta<-signif(x$coef[1], digits=2);#coefficient beta
                          HR <-signif(x$coef[2], digits=2);#exp(beta)
                          HR.confint.lower <- signif(x$conf.int[,"lower .95"], 2)
                          HR.confint.upper <- signif(x$conf.int[,"upper .95"],2)
                          HR <- paste0(HR, " (", 
                                       HR.confint.lower, "-", HR.confint.upper, ")")
                          res<-c(beta, HR, wald.test, p.value)
                          names(res)<-c("beta", "HR (95% CI for HR)", "wald.test(z)", 
                                        "p.value")
                          return(res)
                          return(exp(cbind(coef(x),confint(x))))
                         })
exp.model.out <- t(as.data.frame(exp_results, check.names = FALSE))
as.data.frame(exp.model.out)
```

######################################################################################################
#
#                      Cox Proportional Hazards Models for multiple stressors
#
######################################################################################################

# Start with a full model that contains all of the chronic stressors from both experimental and control conditions
```{r }
# a full model that contains all of our parameters
multiv.mod <- coxph(Surv(Time, Death) ~ Plastic + Light + Food + Temperature + Sym.State, data = surv.dat)
summary(multiv.mod)
```
```{r}
# Perform Log-rank test on our multiv.mod
multiv_diff <- survdiff(Surv(Time, Death) ~ Plastic + Light + Food + Temperature + Sym.State, data = surv.dat)
multiv_diff
```
```{r}
# Let's create a forest plots to show the output of our model
ggforest(multiv.mod, data=surv.dat)
```

# Remove phenotype 
```{r}
# model with chronic stressors only
m1 <- coxph(Surv(Time, Death) ~ Plastic + Light + Food + Temperature, data=surv.dat)
summary(m1)
```
```{r}
# Perform Log-rank test
m1_diff <- survdiff(Surv(Time, Death) ~ Plastic + Light + Food + Temperature,data = surv.dat)
m1_diff
```
```{r ALL STRESSOR FOREST}
# Plot model output 
ggforest(m1, data=surv.dat)
```

# Now we can compare those two models using the likelihood ratio test 
```{r}
anova(multiv.mod, m1, test="LRT")

# a high p value shows that we don't have evidence to say that the full model is better than the reduced model, 
# while a low (and significant p value) tells us that our full model is not made better by dropping the variable and our reduced model is therefore significantly worse than our full model
```

```{r}
# We see that temperature has a very high effect here, what does our data look like when we remove it?
tmp.mod <- coxph(Surv(Time, Death) ~ Plastic + Light + Food, data = surv.dat)
summary(tmp.mod)
ggforest(tmp.mod, data=surv.dat)
```
```{r}
# Is it the same output if we use the exprimental data only?
exp.mod <- coxph(Surv(Time, Death) ~ Plastic + Light + Food, data = exp.data)
summary(exp.mod)
ggforest(exp.mod, data=exp.data)
```

# Finally, we'll break it down by sym state, starting off with our Apo frags

```{r}
# Start with aposymbiotic coral:
apo.mod <- coxph(Surv(Time, Death) ~ Plastic + Light + Food + Temperature, data = apo.dat)
summary(apo.mod)

# plot
ggforest(apo.mod, data=apo.dat)
```
```{r}
# remove temperature (since it's so influential)
apo.mod2 <- coxph(Surv(Time, Death) ~ Plastic + Light + Food , data = apo.dat)
summary(apo.mod2)
```
```{r}
# plot
ggforest(apo.mod2, data = apo.dat)
```

# And the same for our symbiotic fragments
```{r}
sym.mod <- coxph(Surv(Time, Death) ~ Plastic + Light + Food + Temperature, data = sym.dat)
summary(sym.mod)

#plot
ggforest(sym.mod, data=sym.dat)
```
```{r}
# now remove temperature 
sym.mod2 <- coxph(Surv(Time, Death) ~ Plastic + Light + Food , data = sym.dat)
summary(sym.mod2)
```
```{r}
#plot
ggforest(sym.mod2, data = sym.dat)
```

######################################################################################################
#
#                                  Testing the CPH model assumptions
#
######################################################################################################

# We'll start out with our first full model that includes all of our stress parameters and sym state, multiv.mod
```{r}
# Let's test for proportional hazards 
all.hazard <- cox.zph(multiv.mod)
print(all.hazard)

# If the p vale for this test is significant for any of the co-variates (ie the variable p value is less than 0.05) then the hazard associated with that variable is NOT PROPORTIONAL and violates the assumptions
```
```{r}
# Let's plot the assumptions graph
ggcoxzph(all.hazard)
```

# lets test our reduced model that has all of our stress parameters
```{r}
treat.hazard <- cox.zph(m1)
print(treat.hazard) 

# Even though removing sym.state from our CPH model showed that it was worse (in the LRT anova) the p values are all insignificant and indicates that phenotype does not pass the assumptions of the cox proportional hazards model. So, we should not include sym state in our model. 
```
```{r}
# and plot
ggcoxzph(treat.hazard)
# If we see a pattern with the variable over time, that is an indication that the assumptions cannot be met
```
```{r}
# and for our full model without temperature
tmp.hazard <- cox.zph(tmp.mod)
tmp.hazard
```

# Let's take a deeper look at our m1 model (all stress parameters without sym state)
```{r}
# estimated changes in the regression coefficients upon deleting each observation in turn
ggcoxdiagnostics(m1, type = "dfbeta",
                 linear.predictions = FALSE, ggtheme = theme_bw())
```
```{r}
# estimated changes in the coefficients divided by their standard errors
ggcoxdiagnostics(m1, type = "dfbetas",
                 linear.predictions = FALSE, ggtheme = theme_bw())
```
```{r}
#  check outliers by visualizing the deviance residuals. The deviance residual is a normalized transform of the martingale residual. These residuals should be roughtly symmetrically distributed about zero with a standard deviation of 1.
ggcoxdiagnostics(m1, type = "deviance",
                 linear.predictions = FALSE, ggtheme = theme_bw())
# not very symmetrical around 0
```

```{r}
# can we do the same for our single model parameters 
plastic.hazard <- cox.zph(plastic.mod1) # experimental treatments
plastic.hazard # pass

light.hazard <- cox.zph(light.mod1) # experimental treatments
light.hazard # pass

food.hazard <- cox.zph(food.mod1) # experimental treatments
food.hazard # pass 

temp.hazard <- cox.zph(temperature.mod) # all treatments
temp.hazard # pass

symst.hazard <- cox.zph(phen1) # experimental treatments
symst.hazard ## only one that differs over time, not proportional!
```

```{r}
# let's then assess our parameter estimates by sym state
apo.hazard <- cox.zph(apo.mod)
apo.hazard # all good
ggcoxzph(apo.hazard)
```
```{r}
apo2.hazard <- cox.zph(apo.mod2)
apo2.hazard # all good
ggcoxzph(apo2.hazard)
```

```{r}
# sym
sym.hazard <- cox.zph(sym.mod)
sym.hazard #nice 
ggcoxzph(sym.hazard)
```
```{r}
sym2.hazard <- cox.zph(sym.mod2)
sym2.hazard #nice 
ggcoxzph(sym2.hazard)
```

```{r}
treat.col <- palette(brewer.pal(n=9, name = "Set2"))

# Let's visualize a model using ggplot 
fit.1 <- survfit(Surv(Time, Death)~Sym.State, data=surv.dat)
ggsurvplot(fit.1)
ggsurvplot(fit.1, conf.int = TRUE, 
           pval = TRUE, 
           legend.labs = c("Apo", "Sym"),
           legend.title = "Symbiotic State",
           palette = c("black", "brown"))


#ggsurvplot(survfit(all.mod), data=surv.dat, palette = treat.col, ggtheme = theme_minimal())


# food & sym. state
sig.fit <- survfit(Surv(Time, Death) ~ Sym.State+Food, data=surv.dat)
p <- ggsurvplot_facet(sig.fit, surv.dat,
                 facet.by= c("Plastic", "Light"),
                 short.panel.labs = TRUE,
                 palette = "aas",
                 pval = TRUE, 
                 pval.method = TRUE,
                 legend.labs = c("Apo/Fed", "Apo/Starved", "Sym/Fed", "Sym/Starved"))
p
# p + lims(y=c(0.25,1.0))


# facet plot with all 4 variables 
fit <- survfit( Surv(Time, Death) ~ Plastic + Sym.State, data = surv.dat )
ggsurvplot_facet(fit, surv.dat, 
                facet.by = c("Food", "Light"),
                palette = "aaas", 
                pval = TRUE, pval.method = TRUE,
                short.panel.labs = TRUE,
                legend.labs = c("NoPlastic/Apo", "NoPlastic/Sym", "Plastic/Apo", "Plastic/Sym"))

# facet plot not including sym state 
fit2 <- survfit( Surv(Time, Death) ~ Plastic, data = surv.dat )
ggsurvplot_facet(fit2, surv.dat, 
                facet.by = c("Food", "Light"),
                palette = "aaas", 
                pval = TRUE, pval.method = TRUE,
                short.panel.labs = TRUE,
                legend.labs = c("No Plastic", "Plastic"))

# facet plot not including sym state 
fit3 <- survfit( Surv(Time, Death) ~ Light, data = surv.dat )
ggsurvplot_facet(fit3, surv.dat, 
                facet.by = c("Food", "Plastic"),
                palette = "aaas", 
                pval = TRUE, pval.method = TRUE,
                short.panel.labs = TRUE)
    

# facet plot with all 4 variables 
fit4 <- survfit( Surv(Time, Death) ~ Food + Sym.State, data = surv.dat )
ggsurvplot_facet(fit4, surv.dat, 
                facet.by = c("Plastic", "Light"),
                palette = "aaas", 
                pval = TRUE, pval.method = TRUE,
                short.panel.labs = TRUE,
                legend.labs = c("Fed/Apo", "Fed/Sym", "Starved/Apo", "Starved/Sym"))

# facet plot sym state by plastic and light 
fit5 <- survfit( Surv(Time, Death) ~ Food, data = surv.dat )
ggsurvplot_facet(fit5, surv.dat, 
                facet.by = c("Plastic", "Light"),
                palette = "aaas", 
                pval = TRUE, pval.method = TRUE,
                short.panel.labs = TRUE)

```







