---
title: "MM.Growth.RMD"
author: "Sarah Speroff"
date: "2/17/2022"
output:
  pdf_document: default
  html_document: default
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###########################################################################################################################
#
#                                                           Data Prep
#
############################################################################################################################

```{r, message=FALSE}
# Libraries
library(Rmisc)
library(ggplot2)
library(dplyr)
library(tidyr)
library(broom)
library(tidyverse)
library(lme4)
library(lmerTest)
library(emmeans)
library(Rmisc)
library(e1071)
```

```{r}
setwd("/Users/speroffs/Desktop/A.Poculata-Multiple-stressors-thesis-analysis-/Growth")
```
```{r}
#Load data
mass.dat <- read.csv("/Users/speroffs/Desktop/A.Poculata-Multiple-stressors-thesis-analysis-/Growth/Full.ID.Metadata.Frag.metadata.(xrack&control).csv", header = TRUE)

# Subset relevant data
data <- subset(mass.dat, select = c("B.TAG", "Sym.State","Treat.ID", "Temperature", "Plastic", "Light", "Food",
                                    "Initial.Mass..g.", "Final.Mass..g.", "Days.In.Treatment"))

# change column names 
names(data)[names(data) == "Initial.Mass..g."] <- "Initial" 
names(data)[names(data) == "Final.Mass..g."] <- "Final" 
names(data)[names(data) == "Days.In.Treatment"] <- "Time" 
names(data)[names(data) == "Treat.ID"] <- "Treatment" 

# relevel to compare all to control 
data$Treatment <- factor(data$Treatment, levels = c("control", "PDS", "PLS", "NDS", "NLS", "PDF", "PLF", "NDF", "NLF"))

# filter out NAs 
data <- data[complete.cases(data),]
```

###########################################################################################################################
#
#                             Calculate average change in mass over time for all polyps
#
############################################################################################################################

# Data collected in grams, converted to MG and noramlized to time spent in treatment
```{r}
# In new column called Change
data$Change = ((data$Final - data$Initial))*1000/data$Time 
data$Change <- round(data$Change, 3) # round to 3 decimal places
```

# Normality testing on final-initial mass over time:
```{r}
# Let's check the distribution of that growth rate data 
par(mfrow=c(1,2))
hist(data$Change)
qqnorm(data$Change)
qqline(data$Change)
```
```{r}
# check the median-mean skew
skewness(data$Change) # a high skew value
```
```{r}
# let's log transform to see if that helps
data$Logchange = log(data$Change)

par(mfrow=c(1,2))
hist(data$Logchange) # helps our distribution but It produces lots of NaNs...

skewness(data$Logchange)
```
```{r}
# lets square root transform 
data$SQchange = sqrt(data$Change)
hist(data$SQchange)
skewness(data$SQchange) # no skew value, but skill wonky
```
```{r}
# let's try the last option, cube root, to see if that is best 
data$cubechange = (data$Change^1/3)
hist(data$cubechange)
skewness(data$cubechange) # the same position we started with. 
```
# Since we are not running tests or models on this data and are using it to get an idea of which treatment grew or lost the most mass over the duration of the experiment, we will use the raw data as it is. When it comes time to using the "decay" data to determine the effect of treatments and stressors on the rate of change we will reassess the transformations then.

# But for now, let's go back and look at the rate of overall change in each treatment:
```{r}
# Summarize the change in mass by treatment 
mean.sum1 <- summarySE(data, measurevar="Change", groupvars="Treatment") 
temp.sum <- summarySE(data, measurevar="Change", groupvars="Temperature") # what's the difference between ambient and elevated temperature conditions?

```
```{r}
#and plot
change.mass.plot <- 
ggplot(data = mean.sum1, 
       aes(x = Treatment, 
           y = Change, 
           col = Treatment)) +
  
  geom_segment(aes(yend=0,xend=Treatment,size = 2)) +
  geom_hline(yintercept = 0) +
  geom_hline(yintercept = -0.7649045,col="black",lty=2) +
  
  scale_y_continuous(name = "Change Mass (mg*day"^"-1"~")", 
                     breaks=seq(0, -3, by = -0.5)) +
  scale_x_discrete(limits = c("control", "PDS", "PLS", "NDS", "NLS", "PDF", "PLF", "NDF", "NLF"),
                   labels = c("Control", "PDS", "PLS", "NDS", "NLS", "PDF", "PLF", "NDF", "NLF")) +
  scale_color_manual(values = c("#131513","#F53100", "#FF8B47","#C39809","#F6CB3C","#0B8454",
                                "#7BB12F","#3F88C5","#A33461")) +
  
  theme_classic(base_size=13) +
  theme(legend.position = "none",
        legend.title = element_blank(),
        legend.text = element_text(size = 12))

change.mass.plot
```
```{r}
# save plot
ggsave(plot = change.mass.plot,
       filename = "/Users/speroffs/Desktop/A.Poculata-Multiple-stressors-thesis-analysis-/Growth/Plots/change.mass.plot.pdf", width = 8, height = 6, units = "in")
```

```{r}
# Summarize the change in mass by treatment and sym state
mean.sum2 <- summarySE(data, measurevar="Change", groupvars=c("Treatment","Sym.State")) 
```
```{r}
# Plot it 
ggplot(data = mean.sum2, 
       aes(x = Treatment, 
           y = Change, 
           col = Sym.State,)) +
  geom_point(size=1.2) +
  geom_segment(aes(yend=0, xend=Treatment), size=1.2) +

  geom_hline(yintercept = 0, col="black", lty=1) + 
  geom_hline(yintercept = -0.83430380, col="tan", lty=2) +
  geom_hline(yintercept = -0.69474359, col="coral4", lty=2) +
  
  labs(y = "Change in Mass (mg*day"^"-1"~")", y = "Treatment") +
  
  scale_color_manual(labels = c("Aposymbiotic", "Symbiotic"),
                     values = c("tan","coral4")) +
  scale_x_discrete(limits = c("control", "PDS", "PLS", "NDS", "NLS", "PDF", "PLF", "NDF", "NLF"),
                   labels = c("Control", "PDS", "PLS", "NDS", "NLS", "PDF", "PLF", "NDF", "NLF")) +
  
  theme_classic(base_size=13) +
  theme(legend.position = "top",
        legend.title = element_blank(),
        legend.text = element_text(size = 12))
```

###################################################################################################################################
# 
#                                   Calculate change in mass with all time points included 
#
###################################################################################################################################

```{r}
# load data
decaydat <- read.csv("growth - growthtime.csv")

# change column names 
names(decaydat)[names(decaydat) == "Treat.ID"] <- "Treatment" 

# change type 
decaydat$Day <- as.numeric(decaydat$Day)

# relevel
decaydat$Treatment <- factor(decaydat$Treatment, levels = c("Control", "PDS", "PLS", "NDS", "NLS", "PDF", "PLF", "NDF", "NLF"))

# remove NAs
decaydat <- decaydat[complete.cases(decaydat),]

# Convert to mg
decaydat$Massmg <- decaydat$Mass*1000
```

################################################################
#
#   Normality testing and transformations of calculated mass                                            
#
################################################################

```{r}
# Histogram and Quantiles for raw data 
par(mfrow=c(1,2))
hist(decaydat$Massmg, main="Mass(mg)")
qqnorm(decaydat$Massmg, main="Mass (mg)")
qqline(decaydat$Massmg) # not a great fit, seems more like an exponential fit
skewness(decaydat$Massmg) #  positive skew! need some type of transformation
```

```{r}
# Log-transform
decaydat$Logmassmg = log(decaydat$Massmg) #new column with the log mass in grams

par(mfrow=c(1,2))
hist(decaydat$Logmassmg, main="Log Mass (mg)")
qqnorm(decaydat$Logmassmg, main="Log Mass (mg)")
qqline(decaydat$Logmassmg) # a much better histogram and quantiles! 
skewness(decaydat$Logmassmg) # now less than moderately left skewed, much better.

```

```{r}
# lets square root transform 
decaydat$SQmass = sqrt(decaydat$Mass)

par(mfrow=c(1,2))
hist(decaydat$SQmass, main="SQ Mass (mg)")
qqnorm(decaydat$SQmass, main="SQ Mass (mg)")
qqline(decaydat$SQmass) 
skewness(decaydat$SQmass) # better than non-transformed, but more skewed than logmass
```

```{r}
# let's try the last option, cube root, to see if that is best 
decaydat$cubemass = (decaydat$Mass^1/3)

par(mfrow=c(1,2))
hist(decaydat$cubemass, main="Mass^1/3 (mg)")
qqnorm(decaydat$cubemass, main="Mass^1/3 (mg)")
qqline(decaydat$cubemass) 
skewness(decaydat$cubemass) # nope
```

# Since the Log-transformed data helps with normality - we will use the Log-transformed Mass for STATISTICAL testing and reporting. But we will use the untransformed mass for graphing and rate reporting. 
```{r}
# summarize the mass at each interval by treatment 
decay_means <- summarySE(decaydat, measurevar = "Massmg", groupvars = c("Treatment", "Day"))
```
```{r}
# plot 
ggplot(data = decay_means, 
       aes(x = Day, 
           y = Massmg,
           color = Treatment)) +
  geom_point() +
  geom_line() +
  facet_wrap(vars(Treatment)) +
  geom_errorbar(aes(ymax = Massmg+se, ymin = Massmg-se), width = 0.1) +
  ylab("Mass (mg)")+
  xlab("Day") +
  scale_color_manual(name = "", values = c("#131513", "#E02D00", "#FF7E33", 
                                                   "#D6A70A", "#ECDD09", "#075F3C", 
                                                   "#7BB12F", "#35516E", "#3F88C5")) +
  scale_x_continuous(breaks=seq(1, 22, by = 7)) + # change x axis intervals 
  theme(panel.border = element_rect(colour = "black", fill = NA, size = 1)) +
  theme_classic()
```

```{r}
# summarize the mass at each interval by treatment 
decay_means2 <- summarySE(decaydat, measurevar = "Massmg", groupvars = c("Treatment", "Day", "Sym.State"))
```
```{r}
# plot 
ggplot(data = decay_means2, 
       aes(x = Day, 
           y = Massmg, 
           color = Sym.State,
           group=Sym.State)) +
  geom_point() +
  geom_line() +
  facet_wrap(vars(Treatment)) +
  geom_errorbar(aes(ymax = Massmg+se, ymin = Massmg-se), width = .1) +
  ylab("Mass (mg)")+
  xlab("Day") +
  scale_color_manual(name = "Phenotype",
                     labels = c("Apo", "Sym"),
                     values = c("tan","coral4")) +
  scale_x_continuous(breaks=seq(1, 22, by = 7)) + # change x axis intervals 
  theme(panel.border = element_rect(colour = "black", fill = NA, size = 1)) +
  theme_classic()

```

# Since we are most interested in the change of weight over time, let's look at the rate of change over time, or the slope
```{r}
# plot growth of overall polyps

ggplot(data = decay_means, 
       aes(x = Day, 
           y = Massmg,)) +
  geom_point(size=1.2) +
  
  geom_smooth(method="lm", se = FALSE, size=1.2, col="black") +
  geom_errorbar(aes(ymax = Massmg+se, ymin = Massmg-se),width =1, size=1.2) +
  
  labs(x="Day", y = "Mass (mg)") +
  scale_x_continuous(breaks=seq(1, 22, by = 7)) + # change x axis intervals
  facet_wrap(vars(Treatment)) +

  theme_classic(base_size = 13) +
  theme(legend.position = "none",
        legend.title = element_blank(),
        legend.text = element_text(size = 12))


```

```{r}
# plot growth of both phenotypes
ggplot(data = decay_means2, 
       aes(x = Day, 
           y = Massmg,
           color = Sym.State)) +
  geom_point(size=1.2) +
  
  geom_smooth(method="lm", se = FALSE, size=1.2) +
  geom_errorbar(aes(ymax = Massmg+se, ymin = Massmg-se),width =1, size=1.2) +
  
  scale_color_manual(labels = c("Aposymbiotic", "Symbiotic"),
                     values = c("tan","coral4")) +
  
  labs(x="Day", y = "Log Mass (mg)") +
  scale_x_continuous(breaks=seq(1, 22, by = 7)) + # change x axis intervals
  facet_wrap(vars(Treatment)) +

  theme_classic(base_size = 13) +
  theme(legend.position = "top",
        legend.title = element_blank(),
        legend.text = element_text(size = 12))

```

###########################################################################################################################
#
#                            Calculate growth rate for each individual polyp using log transformed data 
#
############################################################################################################################

```{r}
# Run simple LM for the change in log(mass) over time by each individual 

# create grouping factor
by_ID <- group_by(decaydat, B.TAG)
```
```{r}
# Create a df with the parameter coefficients of every treatment Linear model 
ID.slope <- 
  do(by_ID, 
   tidy(                     
     lm(Massmg~Day, data= .)))
ID.slope

# keep the intercept term 
ID.slope <- filter(ID.slope, term == "Day")
#rename the estimate 
names(ID.slope)[names(ID.slope) == "estimate"] <- "Slope"
```

## Run linear model with the log-transformed Mass data for statistical comparisons
```{r}
ID.log.slope <-
  do(by_ID,
     tidy(
       lm(Logmassmg~Day, data=.))) 

ID.log.slope <- filter(ID.log.slope, term == "Day")
names(ID.log.slope)[names(ID.log.slope) == "estimate"] <- "Log.Slope"
names(ID.log.slope)[names(ID.log.slope) == "std.error"] <- "Log.std.error"
names(ID.log.slope)[names(ID.log.slope) == "statistic"] <- "Log.statistic"
names(ID.log.slope)[names(ID.log.slope) == "p.value"] <- "Log.pvalue"

ID.slope <- left_join(ID.slope, ID.log.slope, by = "B.TAG")
```

```{r}
# Add treatment and sym.state values to ID.test dataframe by BTag

# Select variables from original data to keep
decaydat2 <- decaydat %>%
  select(Colony, Sym.State, B.TAG, Treatment, Plastic, Temperature, Food, Light) 

# add variables to new df based on ID
ID.slope <- left_join(decaydat2, ID.slope, by = "B.TAG")

# Remove NAs
ID.slope %>%
  drop_na(Slope)
```
```{r}
# Inspect ID.slope D
ID.slope$Colony <- as.factor(ID.slope$Colony)
ID.slope$Sym.State <- as.factor(ID.slope$Sym.State)
ID.slope$Plastic <- as.factor(ID.slope$Plastic)
ID.slope$Temperature <- as.factor(ID.slope$Temperature)
ID.slope$Food <- as.factor(ID.slope$Food)
ID.slope$Light <- as.factor(ID.slope$Light)
ID.slope$B.TAG <- as.factor(ID.slope$B.TAG)

#remove duplicates (should be 1 value per ID)
duplicated(ID.slope)
ID.slope = ID.slope %>% 
  unique() # should be 480

# Remove NAs
ID.slope <- ID.slope %>%
  drop_na(Slope)  %>% # should be 451 individuals 
  drop_na(Log.Slope)
```

###########################################################################################################################
#
#                                    Question: Does treatment affect growth rate?
#
############################################################################################################################

```{r}
# Run a liner effects mixed model to determine if the treatment affects the growth rate of polyps while taking coral colony into account
treat.lmer <- lmer(Log.Slope~Treatment + (1|Colony), data = ID.slope)
summary(treat.lmer)
```
```{r}
# Run an anova to compare this model to a null model 
null <- lmer(Log.Slope ~ 1 + (1|Colony), data = ID.slope)
anova(null, treat.lmer) 
```
```{r}
# Let's compare the growth rates of corals between the treatments
treat.emmeans <- emmeans(treat.lmer, pairwise ~ Treatment)
summary(treat.emmeans)
```

```{r}
# Summarise
treat.means <- summarySE(ID.slope, measurevar = "Slope", groupvars = "Treatment")
summarySE(ID.slope, measurevar = "Slope", groupvars = "Temperature")

```
```{r}
# plot
all.treat.growth.plot <- 
ggplot(data = treat.means,
       aes(x = Treatment, 
           y = Slope,
           color = Treatment)) +
  geom_point(size=2) +
  geom_errorbar(aes(ymax = Slope+se, ymin = Slope-se), width = 0.2, size=1.2) +
  
  geom_hline(yintercept = -0.8013345, col="black", lty=2) + 
  geom_hline(yintercept = 0, col="black", lty=1) +
  
  scale_y_continuous(name = "Average growth rate (mg*day"^"-1"~")", 
                     limits = c(-3,0), breaks = seq(-3, 0, by = 0.5)) + # set y axis intervals

  labs(x = "Treatment") +
  scale_color_manual(name = "", values = c("#131513","#F53100", "#FF8B47","#C39809","#F6CB3C","#0B8454","#7BB12F","#3F88C5","#A33461")) +
  
  theme_classic(base_size=14) +
  theme(legend.position = "none",
        legend.title = element_blank(),
        legend.text = element_text(size = 12))
all.treat.growth.plot
```
```{r}
# save plot
ggsave(plot = all.treat.growth.plot,
       filename = "/Users/speroffs/Desktop/all.treat.growth.plot.jpg", 
       width = 8, height = 6, units = "in", dpi = 300)

```

###########################################################################################################################
#
#                                    Question: Does temperature affect growth rate?
#
############################################################################################################################

```{r}
# Run a liner effects mixed model to determine if the treatment affects the growth rate of polyps while taking coral colony into account
temp.lmer <- lmer(Log.Slope~Temperature + (1|Colony), data = ID.slope)
summary(temp.lmer)
```
```{r}
# Run an anova to compare this model to a null model 

null <- lmer(Log.Slope ~ 1 + (1|Colony), data = ID.slope)
anova(null, temp.lmer) 
```
```{r}
# Let's compare the growth rates of corals between the treatments
temp.means <- emmeans(temp.lmer, pairwise ~ Temperature)
summary(temp.means)
```

```{r}
# sum
temp.sum <- summarySE(ID.slope, measurevar = "Slope", groupvars = "Temperature")
# plot it 

growth.temp.plot <-
ggplot(data = temp.sum,
       aes(x = Temperature,
           y = Slope)) +
  
  geom_point(size=2) +
  geom_errorbar(aes(ymax = Slope+se, ymin = Slope-se), width = .1, size=1.2) +
  
  scale_x_discrete(name = "", # remove labels for paneling 
                   limits = c("AMBIENT", "ELEVATED"),
                   labels = c("Ambient", "Elevated")) +
  #scale_y_continuous(name = "", 
   #                  limits = c(-0.01, -0.003),
   #                  breaks = seq(-0.01, -0.003, by = 0.001)) +
  
  theme_classic(base_size = 16) +
  theme(legend.position = "none",
        legend.title = element_blank(),
        legend.text = element_text(size = 12))

growth.temp.plot
```
```{r}
# Save plot
ggsave(plot = growth.temp.plot,
       filename = "/Users/speroffs/Desktop/A.Poculata-Multiple-stressors-thesis-analysis-/Growth/Plots/growth.temp.plot.pdf", width = 8, height = 6, units = "in")
```

###########################################################################################################################
#
#                                    Question: Does Plastic affect growth rate?
#
############################################################################################################################

```{r}
# Run a liner effects mixed model to determine if the treatment affects the growth rate of polyps while taking coral colony into account

plastic.lmer <- lmer(Log.Slope~Plastic + (1|Colony), data = ID.slope)
summary(plastic.lmer)
```
```{r}
# Run an anova to compare this model to a null model 
null <- lmer(Log.Slope ~ 1 + (1|Colony), data = ID.slope)
anova(null, plastic.lmer) 
```
```{r}
# Let's compare the growth rates of corals between the treatments
plastic.means <- emmeans(plastic.lmer, pairwise ~ Plastic)
summary(plastic.means)
```

```{r}
# sum
plastic.sum <- summarySE(ID.slope, measurevar = "Slope", groupvars = "Plastic")

# plot it 
growth.plastic.plot <- 
ggplot(data = plastic.sum,
       aes(x = Plastic,
           y = Slope)) +
  geom_point(size=2) +
  geom_errorbar(aes(ymax = Slope+se, ymin = Slope-se), width = .1, size=1.2) +
  
  scale_x_discrete(name = "", # remove labels for paneling 
                   limits = c("NOPLASTIC", "PLASTIC"),
                   labels = c("No Plastic", "Plastic")) +
 # scale_y_continuous(name = "", 
 #                    limits = c(-0.01, -0.003),
 #                    breaks = seq(-0.01, -0.003, by = 0.001)) +

  
  theme_classic(base_size = 16) +
  theme(legend.position = "none",
        legend.title = element_blank(),
        legend.text = element_text(size = 12))
growth.plastic.plot
```
```{r}
# Save plot
ggsave(plot = growth.plastic.plot,
       filename = "/Users/speroffs/Desktop/A.Poculata-Multiple-stressors-thesis-analysis-/Growth/Plots/growth.plastic.plot.pdf", width = 8, height = 6, units = "in")
```

###########################################################################################################################
#
#                                    Question: Does Light affect growth rate?
#
############################################################################################################################

```{r}
# Run a liner effects mixed model to determine if the treatment affects the growth rate of polyps while taking coral colony into account

light.lmer <- lmer(Log.Slope~Light + (1|Colony), data = ID.slope)
summary(light.lmer)
```
```{r}
# Run an anova to compare this model to a null model 

null <- lmer(Log.Slope ~ 1 + (1|Colony), data = ID.slope)
anova(null, light.lmer) 
```
```{r}
# Let's compare the growth rates of corals between the treatments
light.means <- emmeans(light.lmer, pairwise ~ Light)
summary(light.means)
```

```{r}
# sum
light.sum <- summarySE(ID.slope, measurevar = "Slope", groupvars = "Light")

# plot it 
growth.light.plot <- 
ggplot(data = light.sum,
       aes(x = Light,
           y = Slope)) +
  geom_point(size=2) +
  geom_errorbar(aes(ymax = Slope+se, ymin = Slope-se), width = .1, size=1.2) +
  
  scale_x_discrete(name = "", # remove labels for paneling 
                   limits = c("DARK", "LIGHT"),
                   labels = c("Dark ", "Light")) +
  #scale_y_continuous(name = "", 
  #                   limits = c(-0.01, -0.003),
  #                   breaks = seq(-0.01, -0.003, by = 0.001)) +

  theme_classic(base_size = 16) +
  theme(legend.position = "none",
        legend.title = element_blank(),
        legend.text = element_text(size = 12))
growth.light.plot
```
```{r}
# Save plot
ggsave(plot = growth.light.plot,
       filename = "/Users/speroffs/Desktop/A.Poculata-Multiple-stressors-thesis-analysis-/Growth/Plots/growth.light.plot.pdf", width = 8, height = 6, units = "in")

```

###########################################################################################################################
#
#                                    Question: Does Food affect growth rate?
#
############################################################################################################################

```{r}
# Run a liner effects mixed model to determine if the treatment affects the growth rate of polyps while taking coral colony into account

food.lmer <- lmer(Log.Slope~Food + (1|Colony), data = ID.slope)
summary(food.lmer)
```
```{r}
# Run an anova to compare this model to a null model 

null <- lmer(Log.Slope ~ 1 + (1|Colony), data = ID.slope)
anova(null, food.lmer) 
```
```{r}
# Let's compare the growth rates of corals between the treatments
food.means <- emmeans(food.lmer, pairwise ~ Food)
summary(food.means)
```

```{r}
# sum
food.sum <- summarySE(ID.slope, measurevar = "Slope", groupvars = "Food")

# plot it 
growth.food.plot <-
ggplot(data = food.sum,
       aes(x = Food,
           y = Slope)) +
  geom_point(size=2) +
  geom_errorbar(aes(ymax = Slope+se, ymin = Slope-se), width = .1, size=1.2) +
  
  scale_x_discrete(name = "", # remove labels for paneling 
                   limits = c("FED", "STARVED"),
                   labels = c("Fed ", "Starved")) +
  #scale_y_continuous(name = "", 
  #                   limits = c(-0.01, -0.003),
  #                   breaks = seq(-0.01, -0.003, by = 0.001)) +

  theme_classic(base_size = 16) +
  theme(legend.position = "none",
        legend.title = element_blank(),
        legend.text = element_text(size = 12))
growth.food.plot
```
```{r}
# Save plot
ggsave(plot = growth.food.plot,
       filename = "/Users/speroffs/Desktop/A.Poculata-Multiple-stressors-thesis-analysis-/Growth/Plots/growth.food.plot.pdf", width = 8, height = 6, units = "in")

```

###########################################################################################################################
#
#                                    Question: Does Phenotype affect growth rate?
#
############################################################################################################################

```{r}
# Run a liner effects mixed model to determine if the treatment affects the growth rate of polyps while taking coral colony into account

phen.lmer <- lmer(Log.Slope~Sym.State + (1|Colony), data = ID.slope)
summary(phen.lmer)
```
```{r}
# Run an anova to compare this model to a null model 

null <- lmer(Log.Slope ~ 1 + (1|Colony), data = ID.slope)
anova(null, phen.lmer) 
```
```{r}
# Let's compare the growth rates of corals between the treatments
phen.means <- emmeans(phen.lmer, pairwise ~ Sym.State)
summary(phen.means)
```

```{r}
# sum
phen.sum <- summarySE(ID.slope, measurevar = "Slope", groupvars = "Sym.State")
# plot it 
ggplot(data = phen.sum,
       aes(x = Sym.State,
           y = Slope, 
           color = Sym.State)) +
  geom_point() +
  geom_errorbar(aes(ymax = Slope+se, ymin = Slope-se), width = .1) +
  scale_color_manual(values = c("tan","coral4")) +
  labs(y = "Growth rate log(mg*day"^"-1"~")", 
       x = "") +
  theme_classic()
```

###########################################################################################################################
#
#                                                     Mixed effects models:
#
############################################################################################################################

```{r}
# Null model 
m0 <- lmer(Log.Slope ~ 1 + (1|Colony), data=ID.slope)
summary(m0)
```

# Start with our 4 stressors
```{r}
# Full model
m1 <- lmer(Log.Slope ~ Temperature + Plastic + Light + Food + (1|Colony), data = ID.slope)
summary(m1)
```
```{r}
# Drop plastic
m2 <- lmer(Log.Slope ~ Temperature + Light + Food + (1|Colony), data = ID.slope)
summary(m2) # all significant

anova(m0, m1, m2) # m2
```

# Consider phenotype
```{r}
m3 <- lmer(Log.Slope ~ Temperature + Plastic + Light + Food + Sym.State + (1|Colony), data = ID.slope)
summary(m3)
```
```{r}
# drop plastic
m4 <- lmer(Log.Slope ~ Temperature + Light + Food + Sym.State + (1|Colony), data = ID.slope)
summary(m4) # sym not significant, if we drop we're at the same model as m2
```

# Let's see if there are any interactions between these significant factors 
```{r}
m5 <- lmer(Log.Slope ~ Temperature + Light + Food + Light*Food + (1|Colony), data = ID.slope)
summary(m5)
```
```{r}
anova(m0, m1, m2, m3, m4, m5)
```

# Let's check our m2 model for model assumptions
```{r}
library(performance)
check_model(m2) # actually looks alright
```

```{r}
#libraries required
library(lme4)
library(sjPlot)

# plot model outputs 
tab_model(m1, m2,             # models to table
          show.intercept = TRUE,  # show intercept value
          p.val = "kr",           # p value method for lmer models, kenward-rogers
          p.style = "numeric",
          show.df = FALSE,        # degrees freedom
          show.aic = TRUE,        # AIC score
          show.icc = FALSE,       # contrast correlation for mm  
          show.r2 = FALSE,
          show.stat = TRUE,
          show.ngroups = FALSE,
          show.loglik = TRUE,
          show.ci = FALSE,
          digits = 4,
          show.re.var = FALSE,
          title = "")
          
```

# Let's plot our model estimates
```{r}
library(insight)
library(sjstats)
library(sjPlot)
library(effects)
```
```{r}
m <- lmer(Log.Slope ~ Temperature + Light + Food + (1|Colony), data = ID.slope)
```
```{r}
# plot light
l=plot_model(m, 'pred', ci.lvl = 0.95, terms="Light")

l +
  theme_classic() +
xlab("Light")+
  ylab("Growth rate log(mass) (mg*day^-1)")+
  ggtitle("")+
  theme(axis.title.x = element_text(color="black", size=14, face="bold"),
axis.title.y = element_text(color="black", size=14, face="bold"))+
  theme(axis.text.x = element_text(size=14),
          axis.text.y = element_text(size=14))
```
```{r}
# plot temperature
t=plot_model(m, 'pred', ci.lvl = 0.95, terms="Temperature")

t +
  theme_classic() +
xlab("Temperature")+
  ylab("Growth rate log(mass) (mg*day^-1)")+
  ggtitle("")+
  theme(axis.title.x = element_text(color="black", size=14, face="bold"),
axis.title.y = element_text(color="black", size=14, face="bold"))+
  theme(axis.text.x = element_text(size=14),
          axis.text.y = element_text(size=14))
```
```{r}
# plot food
f=plot_model(m, 'pred', ci.lvl = 0.95, terms="Food")

f +
  theme_classic() +
xlab("Food")+
  ylab("Growth rate log(mass) (mg*day^-1)")+
  ggtitle("")+
  theme(axis.title.x = element_text(color="black", size=14, face="bold"),
axis.title.y = element_text(color="black", size=14, face="bold"))+
  theme(axis.text.x = element_text(size=14),
          axis.text.y = element_text(size=14))
```

```{r}
# Plot interactions  
library(ggeffects)

# Interaction plot model 
int.m<-lmer(Log.Slope ~ Temperature + Light + Food + (1|Colony), data = ID.slope)
summary(int.m)

dat2 <- ggemmeans(int.m, terms = c("Temperature", "Light", "Food")) #

ggplot(dat2, aes(x,predicted, colour = group)) +
  geom_point() +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.3) +
  labs(x = "Temperature", y = "Rate O2 Consumption (umol*min-1*g-1)",
       color = "Light") +
  facet_grid(~facet) +
  ggtitle("") +
  theme_classic() +
  theme(axis.title.x = element_text(color="black", size=12),
axis.title.y = element_text(color="black", size=12))+
  theme(axis.text.x = element_text(size=12),
          axis.text.y = element_text(size=12)) +
  scale_color_manual(values=c("black", "goldenrod"))
```


###########################################################################################################################
#
#                                                 Format for merged metric datasheet
#
############################################################################################################################

```{r}
# load in data that includes mass on day 22 for NDF and NLF treatments (not used for analysis)
all.mass.dat <- read.csv("~/Desktop/A.Poculata-Multiple-stressors-thesis-analysis-/Growth/growth.over.time.withNDF.NLF.day22.data.csv", header = TRUE)

# switch to wide format
wide.mass = all.mass.dat %>%
  select(B.TAG, Day, Mass) %>%
  spread(Day, Mass)
```
```{r}
# load in all.data.merged
all.data.merged <- read.csv("~/Desktop/A.Poculata-Multiple-stressors-thesis-analysis-/All.data.merged.csv", header=TRUE)
names(all.data.merged)
```

```{r}
# edit names of wide.mass
names(wide.mass)[names(wide.mass) == "B.TAG"] <- "BTag" 
names(wide.mass)[names(wide.mass) == "1"] <- "MassG.day0" 
names(wide.mass)[names(wide.mass) == "8"] <- "MassG.day8" 
names(wide.mass)[names(wide.mass) == "15"] <- "MassG.day15" 
names(wide.mass)[names(wide.mass) == "22"] <- "MassG.day22" 
```

```{r}
# add wide.mass to all.data.merged
all.data.merged <- left_join(all.data.merged, wide.mass, by = "BTag")
```
```{r}
# save merged data as a csv
path <- "/Users/speroffs/Desktop/A.Poculata-Multiple-stressors-thesis-analysis-"
write.csv(all.data.merged, file.path(path, "All.data.merged.csv"))
```










